Ornaments as indicators of social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-08-14 analysis/paper/paper.Rmd
update figure code after it disappeared :$

---
title: "Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan"
author:
  - Li-Ying Wang
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

# Excavations at Kiwulan in northeastern Taiwan

```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", out.width= "400px"}
# need to add the locations of Keelung, Tamsui, and hoping dao
library(here)
knitr::include_graphics(here("analysis", "figures", "kiwulan-location-map.png"))
```

@@

```{r tidy-data, eval = TRUE}
# summarise the ornaments by layers
# L5 and L6 represent pre-European contact period, 
# L4 was the time of European contact, and 
# L2 and L1 was the period of Chinese contact.

ornaments_period <- 
kwl_upper %>% 
  filter(!is.na(`6-layers`)) %>% 
  mutate(period = case_when(
    `6-layers` %in% 1:2 ~ "Chinese Contact",
    `6-layers` == 4 ~ "European Contact",
    `6-layers` %in% 5:6 ~ "Before European Contact",
     TRUE ~ "other"
  )) %>% 
  filter(period != "other") %>% 
  mutate(period = factor(period,
                       level = c("Before European Contact", 
                                 "European Contact", 
                                 "Chinese Contact"))) %>% 
  mutate(Length = as.numeric(`Length(mm)`),
         Thick = as.numeric(`Thick(mm)`),
         Width = as.numeric(`Width(mm)`),
         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
  rowwise() %>% 
  mutate(Perforation_average =
           mean(c(`Perforation1`,`Perforation2(mm)`), 
                na.rm = TRUE))
```

```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number. Each square is 4 x 4 m", out.width= "70%", echo=FALSE}
# add to the map and scale and north arrow. Check it out after knitting 
knitr::include_graphics(here("analysis", "figures", "KWL-excavation-map.png"))
```

@@

```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan", out.width='60%'}
# need to reorder, try https://forcats.tidyverse.org/reference/index.html
# reorder factor for categories (by frequency) and period (manually)
# total number of ornaments for each categories

ornaments_count <-
  ornaments_period %>% 
  filter(Categories != "Unknown Metal") %>%
  ungroup() %>% # need to upgroup because of rowwise or factor in previous chunk?
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  group_by(Categories) %>%
  count() %>% 
  ungroup() %>% 
  mutate(Categories = fct_reorder(Categories, n))

ggplot(ornaments_count,
       aes(Categories, n)) +
  geom_col() +
  coord_flip() +
  labs(y = "Frequency",
       x = "Categories") +
  theme_minimal(base_size = 14)
```

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "glass bead", "medium bead", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type),
         Type = ifelse(Type == "large thin string", "thin large", Type))

ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>%
  filter(!Type %in% c("other",
                      "unclassified")) %>% 
  arrange(Categories, desc(`European Contact`))

# making table
knitr::kable(ornaments_shape,
             caption = "Ornament subtype frequencies at Kiwulan")

# some basic stat
sam_post_hole <- nrow(ornaments_period)
total_post_hole <- nrow(kwl_upper)
sample_per <- round(sam_post_hole/total_post_hole*100, 1)
```

```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class. A: agate beads, B: bells, C: glass beads and golden beads, D: metal rings. Photographs are presented in the same order as those subtypes in the table but from left to right instead. The photographs of B, C, D classes were from original excavation report (Chen 2007).", out.width= "400px", echo=FALSE}
knitr::include_graphics(here("analysis", "figures", "ornament.png"))
```

@@

```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods."}
### delete one plot

# categories of ornaments by period
ornaments_count <-
  ornaments_period %>% 
  filter(!is.na(`6-layers`),
         Categories != "Unknown Metal") %>%
  ungroup() %>% 
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  mutate(Categories = fct_infreq(Categories)) %>% 
  mutate(Categories = fct_rev(Categories)) %>% 
  group_by(period, Categories) %>%
  count()

# histogram facetting for three periods
ggplot(ornaments_count) +
  geom_col(aes(Categories, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(x = "categories", y = "Frequency") +
  theme_minimal(base_size = 12)
```

```{r chi-square-testing}
# prepare contingency table for chi sqaure
ornaments_shape_chi <-
  ornaments_period_clean %>% 
  group_by(period, Categories) %>% 
  tally() %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>% 
  spread(Categories, n)

# compute chi sqaure
chisq_test_result <- chisq.test(ornaments_shape_chi[ , -1])
chisq_test_result_stat <- round(chisq_test_result$statistic, 2)
chisq_test_result_df <- round(chisq_test_result$parameter, 2)
chisq_test_result_p_value <- (chisq_test_result$p.value)
```

@@

Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequency for the major class of ornaments between different time periods at Kiwulan. The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = `r chisq_test_result_p_value`) shows that there is a non-random difference between the three time periods. It indicates that most ornament types were present before European contact. Ornament frequencies reached a peak during European contact and then dropped during the Chinese contact period, especially for golden beads. This trend can be also seen on other ornaments including agate beads, metal rings, and bells. However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European contact period and a further decrease in the Chinese contact period.

@@

```{r agate-bead-fre-plot}
# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Agate bead',
         Type != 'other')

agate_bead_shape_plot <- 
ggplot(agate_bead_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Agate bead") +
  theme_minimal(base_size = 14) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r metal-fre-plot}
# plot for different shape of metal ring by period
ring_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal ring',
         !Type %in% c('large thick string',
                       'small thin string')) %>% 
  ungroup() %>% 
  mutate(Type = fct_drop(Type)) %>% 
  mutate(Type = fct_infreq(Type)) %>% 
  mutate(Type = fct_rev(Type))

ring_shape_plot <-
ggplot(ring_shape) +
  geom_col(aes(Type, n)) +
  coord_flip() +
  labs(y = "",
       x = "Metal ring") +
  facet_wrap(~ period) +
  theme_minimal(base_size = 14) 
```

```{r bell-fre-plot}
# plot for different shape of metal bell by period
bell_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Bell',
         Type != 'unclassified')

bell_shape_plot <-
ggplot(bell_shape) +
  geom_col(aes(reorder(Type,n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Bell") +
  theme_minimal(base_size = 14) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r glass-bead-fre-plot}
# plot for different shape of glass bead by period
glass_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Glass bead')

glass_bead_plot <-
ggplot(glass_bead_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Glass bead") +
  theme_minimal(base_size = 14) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across contact periods for metal rings, agate beads, bells, and glass beads."}
library(cowplot)
plot_grid(ring_shape_plot,
          agate_bead_shape_plot, 
          bell_shape_plot, 
          glass_bead_plot,
          ncol=1,
          align = "h",
          axis = "lr",
          rel_widths = 1,
          rel_heights = c(1,1,0.7,0.62))
```

```{r regression}
# tidy up the data for correlation testing
ornaments_diversity_test <-
  ornaments_shape_count %>%
  group_by(period) %>% 
  mutate(sam_size = sum(n)) %>% 
  add_count() %>% 
  select(period, n, sam_size) %>% 
  distinct()

ornaments_diversity_test_2 <-
  ornaments_shape_count %>%
  unite(diversity, c(period, Categories), sep = "-", remove = FALSE) %>% 
  group_by(diversity) %>% 
  mutate(sam_size = sum(n)) %>% 
  add_count() %>% 
  select(diversity, n, sam_size) %>% 
  separate(diversity, c("period","subtype"), sep = "-", remove = FALSE) %>% 
  distinct()

# spearman test
spearman_result <- 
cor.test( ~ n + sam_size, 
         data = ornaments_diversity_test_2,
         method = "spearman",
         continuity = FALSE,
         conf.level = 0.95)

spearman_result_stat <- round(spearman_result$statistic, 2)
spearman_result_rho <- unname(round(spearman_result$estimate, 2))
spearman_result_p_value <- round(spearman_result$p.value, 2)
```

The distribution of frequency for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether). Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p-value = `r spearman_result_p_value`). This indicates that the increases in diversity can be explained by the effects of culture contact instead of effects of sample size. According to the frequency, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period. The greater varieties for agate beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the Europeans. In contrast, copper bells and glass beads have less variety, but glass beads have a larger number because they were usually found in a cluster that indicates the original use as strings or necklaces. The common shape for copper bells is the large shape with a wide variety of human faces as a motif, while the common shape for glass beads is a bead with length less than 1 cm that are often identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].

```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
library(sf)
# read in the spatial data
AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))

# tidy spatial data for 40 sampling units
sample_units <-
  AD_data %>%
  semi_join(units_40, by = c (`坑號` = "the_sq"))

# plot spatial pattern for all ornaments
ggplot() +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join, 
          aes(geometry = geometry, 
              fill =n), 
          alpha = 0.9) +
  facet_wrap(~ period) +
  labs(fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))
```

```{r, plot-spatial-pattern-each, fig.cap= "Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}

# filter golden bead
ornaments_period_join_gold <-
  ornaments_period_join %>% 
  filter(Categories %in% 'Golden bead')

# plot by period
dis_golden <-
ggplot(ornaments_period_join_gold) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_gold, 
          aes(geometry = geometry,
              fill = n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(title = "Golden Beads",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))

# filter glass bead
ornaments_period_join_gold <-
  ornaments_period_join %>% 
  filter(Categories %in% 'Glass bead')

# plot by period
dis_glass <-
  ggplot(ornaments_period_join_gold) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_gold, 
          aes(geometry = geometry,
              fill = n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(title = "Glass Beads",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))

# filter agate bead
ornaments_period_join_gold <-
  ornaments_period_join %>% 
  filter(Categories %in% 'Agate bead')

# plot by period
dis_agate <-
ggplot(ornaments_period_join_gold) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_gold, 
          aes(geometry = geometry, 
              fill = n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(title = "Agate Beads",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))

plot_grid(dis_golden,
          dis_glass,
          dis_agate, 
          align = "v",
          nrow = 3)
```

@@

```{r point-pattern-data-tidy}
# get centroid and make a list of three periods
ornaments_period_join_cen_all <-
  ornaments_period_join %>% 
  mutate(lon = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[1]])),
         lat = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[2]])))

ornaments_period_join_cen <- 
  ornaments_period_join_cen_all %>% 
  group_by(period) %>% 
  group_split() %>% 
  set_names(c("Pre-European",
              "European",
              "Chinese"))
```

```{r transformation-sptial-data}
# transform data to PPP format and sp SpatialPoints
library(spatstat)
library(maptools)
library(raster)

#store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points, by period

# use common range for each period
lon_all <- ornaments_period_join_cen_all$lon
lat_all <- ornaments_period_join_cen_all$lat
xrange <- range(lon_all, na.rm=T)
yrange <- range(lat_all, na.rm=T)

# create a function to apply to each period
density_analysis <- function(x, ...) {
  # create ppp and sp(spacial points) format
lon_tmp <- x$lon
lat_tmp <- x$lat
ornaments_point_ppp_tmp <- ppp(lon_tmp, lat_tmp, xrange, yrange)
ornaments_point_sp_tmp <- as(ornaments_point_ppp_tmp, "SpatialPoints")

>>>>

---
title: "Ornaments as indicators of social change before and after European contact at Kiwulan, Northeastern Taiwan"
author:
  - Li-Ying Wang
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

# Excavations at Kiwulan in northeastern Taiwan
```{r, KWL-map, fig.cap = "Map showing the location of Kiwulan, and other places in northern Taiwan named in the text", fig.width=image_width_2_col_inch}
# need to add the locations of Keelung, Tamsui, and hoping dao
library(here)
knitr::include_graphics(here("analysis", "figures", "kiwulan-location-map.png"))
```

@@

```{r tidy-data, eval = TRUE}
# summarise the ornaments by layers
# L5 and L6 represent pre-European contact period, 
# L4 was the time of European contact, and 
# L2 and L1 was the period of Chinese contact. 
ornaments_period <- 
kwl_upper %>% 
  filter(!is.na(`6-layers`)) %>% 
  mutate(period = case_when(
    `6-layers` %in% 1:2 ~ "Chinese Contact",
    `6-layers` == 4 ~ "European Contact",
    `6-layers` %in% 5:6 ~ "Before European Contact",
     TRUE ~ "other"
  )) %>% 
  filter(period != "other") %>% 
  mutate(period = factor(period,
                       level = c("Before European Contact", 
                                 "European Contact", 
                                 "Chinese Contact"))) %>% 
  mutate(Length = as.numeric(`Length(mm)`),
         Thick = as.numeric(`Thick(mm)`),
         Width = as.numeric(`Width(mm)`),
         Perforation1 = as.numeric(`Perforation1(mm)`)) %>% 
  rowwise() %>% 
  mutate(Perforation_average =
           mean(c(`Perforation1`,`Perforation2(mm)`), 
                na.rm = TRUE))
```

```{r, KWL-sam-area, fig.cap = "Map showing the largest section of excavation areas at Kiwulan, and the distribution of forty squares sampled in this paper presented in red with square ID number. Each square is 4 x 4 m", fig.width=image_width_2_col_inch}
# add to the map and scale and north arrow. Check it out after knitting 
library(sf)
library(cartography)
# read in the spatial data
burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
location <- invisible(st_read(here("analysis", "data", "raw_data", "location.shp"), quiet = TRUE))
units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))

# tidy spatial data
# 40 sampling units
sample_units <-
  AD_data %>%
  semi_join(units_40, by = c (`坑號` = "the_sq"))

# mapping sampling area
KWL_excavation_map <- 
ggplot(sample_units) +
  geom_sf(data = AD_data, fill = NA) +
  geom_sf(color = 'red') +
  geom_sf_text(aes(label = `坑號`), size = 2.5, alpha = 0.9) +
  theme_minimal() +
  theme_void()

#1. Need to say that this is a map of excavation areas at Kiwulan, because the reader cannot see the site name anywhere in this figure.
#2. Need to add in here that each square is 4 x 4 m
#3. and say which way north is. Or add to the map and scale and north arrow.

ggsave(plot = KWL_excavation_map, 
       here("analysis", "figures", "KWL-excavation-map.png"), width = 190, units = "mm")

knitr::include_graphics(here("analysis", "figures", "KWL-excavation-map.png"))
```

@@

```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan", fig.width=image_width_1_col_inch}
# need to reorder, try https://forcats.tidyverse.org/reference/index.html
# reorder factor for categories (by frequency) and period (manually)
# total number of ornaments for each categories 
ornaments_count <-
  ornaments_period %>% 
  filter(Categories != "Unknown Metal") %>%
  ungroup() %>% # need to upgroup because of rowwise or factor in previous chunk?
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  group_by(Categories) %>%
  count() %>% 
  ungroup() %>% 
  mutate(Categories = fct_reorder(Categories, n))
plot_ornaments_count <- 
ggplot(ornaments_count,
       aes(Categories, n)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(limits = c(0, 160)) +
  labs(y = "Frequency",
       x = "Categories") +
  theme_minimal(base_size = 10)
ggsave(plot = plot_ornaments_count,
       here("analysis", "figures",
       "plot-ornaments-count.png"),
       width = 90, 
       height = 40,
       units = "mm")
knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-ornaments-count.png"))
```

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "glass bead", "medium bead", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type),
         Type = ifelse(Type == "large thin string", "thin large", Type))
ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>%
  filter(!Type %in% c("other",
                      "unclassified")) %>% 
  arrange(Categories, desc(`European Contact`))
# making table
knitr::kable(ornaments_shape,
             caption = "Ornament subtype frequencies at Kiwulan")
# some basic stat
sam_post_hole <- nrow(ornaments_period)
total_post_hole <- nrow(kwl_upper)
sample_per <- round(sam_post_hole/total_post_hole*100, 1)
```

```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class. A: agate beads, B: bells, C: glass beads and golden beads, D: metal rings. Photographs are presented in the same order as those subtypes in the table but from left to right instead. The photographs of B, C, D classes were from original excavation report (Chen 2007).", echo=FALSE}
knitr::include_graphics(here("analysis", "figures", "ornament.png"))
```

@@

```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods."}
### delete one plot
# categories of ornaments by period
ornaments_count <-
  ornaments_period %>% 
  filter(!is.na(`6-layers`),
         Categories != "Unknown Metal") %>%
  ungroup() %>% 
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  mutate(Categories = fct_infreq(Categories)) %>% 
  mutate(Categories = fct_rev(Categories)) %>% 
  group_by(period, Categories) %>%
  count()
# histogram facetting for three periods
plot_ornaments_count_three_periods <- 
ggplot(ornaments_count) +
  geom_col(aes(Categories, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(x = "Categories",
       y = "Frequency") +
  theme_minimal(base_size = 12)

ggsave(plot = plot_ornaments_count_three_periods,
       here("analysis", "figures",
       "plot-ornaments-count-three-periods.png"),
       width = 190, 
       height = 80,
       units = "mm")

knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-ornaments-count-three-periods.png"))
```

```{r chi-square-testing}
# prepare contingency table for chi sqaure
ornaments_shape_chi <-
  ornaments_period_clean %>% 
  group_by(period, Categories) %>% 
  tally() %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>% 
  spread(Categories, n) 
# compute chi sqaure
chisq_test_result <- chisq.test(ornaments_shape_chi[ , -1])
chisq_test_result_stat <- round(chisq_test_result$statistic, 2)
chisq_test_result_df <- round(chisq_test_result$parameter, 2)
chisq_test_result_p_value <- (chisq_test_result$p.value)
```

@@

Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequency for the major class of ornaments between different time periods at Kiwulan. The result of chi-square test for the number of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = $`r chisq_test_result_p_value`$) shows that there is a non-random difference between the three time periods. It indicates that most ornament types were present before European contact. Ornament frequencies reached a peak during European contact and then dropped during the Chinese contact period, especially for golden beads. This trend can be also seen on other ornaments including agate beads, metal rings, and bells. However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European contact period and a further decrease in the Chinese contact period.

@@

```{r agate-bead-fre-plot}
# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Agate bead',
         Type != 'other')
base_size_multi_plot <- 10
agate_bead_shape_plot <- 
ggplot(agate_bead_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Agate bead") +
  theme_minimal(base_size = base_size_multi_plot) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r metal-fre-plot}
# plot for different shape of metal ring by period
ring_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal ring',
         !Type %in% c('large thick string',
                       'small thin string')) %>% 
  ungroup() %>% 
  mutate(Type = fct_drop(Type)) %>% 
  mutate(Type = fct_infreq(Type)) %>% 
  mutate(Type = fct_rev(Type)) 
ring_shape_plot <-
ggplot(ring_shape) +
  geom_col(aes(Type, n)) +
  coord_flip() +
  labs(y = "",
       x = "Metal ring") +
  facet_wrap(~ period) +
  theme_minimal(base_size = base_size_multi_plot) 
```

```{r bell-fre-plot}
# plot for different shape of metal bell by period
bell_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Bell',
         Type != 'unclassified')
bell_shape_plot <-
ggplot(bell_shape) +
  geom_col(aes(reorder(Type,n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Bell") +
  theme_minimal(base_size = base_size_multi_plot) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r glass-bead-fre-plot}
# plot for different shape of glass bead by period
glass_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Glass bead')
glass_bead_plot <-
ggplot(glass_bead_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Glass bead") +
  theme_minimal(base_size = base_size_multi_plot) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

```{r plot-altogether, fig.cap= "Frequency of ornament subtypes showing the changes in frequency across contact periods for metal rings, agate beads, bells, and glass beads."}
library(cowplot)
plot_multi_types_by_period <- 
plot_grid(ring_shape_plot,
          agate_bead_shape_plot, 
          bell_shape_plot, 
          glass_bead_plot,
          ncol=1,
          align = "h",
          axis = "lr",
          rel_widths = 1,
          rel_heights = c(1,1,0.7,0.62))
ggsave(plot = plot_multi_types_by_period,
       here("analysis", "figures",
       "plot-multi-types-by-period.png"),
       width = 190, 
       height = 150,
       units = "mm")
knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-multi-types-by-period.png"))
```

```{r regression}
# tidy up the data for correlation testing
ornaments_diversity_test <-
  ornaments_shape_count %>%
  group_by(period) %>% 
  mutate(sam_size = sum(n)) %>% 
  add_count() %>% 
  select(period, n, sam_size) %>% 
  distinct()
ornaments_diversity_test_2 <-
  ornaments_shape_count %>%
  unite(diversity, c(period, Categories), sep = "-", remove = FALSE) %>% 
  group_by(diversity) %>% 
  mutate(sam_size = sum(n)) %>% 
  add_count() %>% 
  select(diversity, n, sam_size) %>% 
  separate(diversity, c("period","subtype"), sep = "-", remove = FALSE) %>% 
  distinct()
# spearman test
spearman_result <- 
cor.test( ~ n + sam_size, 
         data = ornaments_diversity_test_2,
         method = "spearman",
         continuity = FALSE,
         conf.level = 0.95)
spearman_result_stat <- round(spearman_result$statistic, 2)
spearman_result_rho <- unname(round(spearman_result$estimate, 2))
spearman_result_p_value <- round(spearman_result$p.value, 2)
```

The distribution of frequency for subtypes in each major class are presented in Figure \@ref(fig:plot-altogether). Spearman's correlation test shows that there is no significant relationship between diversity of subtypes and sample size (S = `r spearman_result_stat`, rho = `r spearman_result_rho`, p = $`r spearman_result_p_value`$). This indicates that the increases in diversity can be explained by the effects of culture contact instead of effects of sample size. According to the frequency, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period. The greater varieties for agate beads and metal rings might indicate multiple origins due to participation in global trade networks stimulated by the Europeans. In contrast, copper bells and glass beads have less variety, but glass beads have a larger number because they were usually found in a cluster that indicates the original use as strings or necklaces. The common shape for copper bells is the large shape with a wide variety of human faces as a motif, while the common shape for glass beads is a bead with length less than 1 cm that are often identified as Indo-Pacific beads, the most common type that had been widespread in Southeast Asian sites since 300 BC [@Francis2002].

```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of all class of ornament by time periods"}
library(sf)

# read in the spatial data
AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
# tidy spatial data for 40 sampling units
sample_units <-
  AD_data %>%
  semi_join(units_40, by = c (`坑號` = "the_sq"))
# join two dataset to get coordinate

ornaments_period_join <-
  ornaments_period %>%
  mutate(period = as.factor(period)) %>% # seems need this line first
  group_by(Pit_No, period) %>%
  count(Categories) %>%
  inner_join(AD_data, by = c("Pit_No" = "坑號"))
# plot spatial pattern for all ornaments

plot_spatial_distribution_all_ornaments_by_period <- 
ggplot() +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join, 
          aes(geometry = geometry, 
              fill =n), 
          alpha = 0.9) +
  facet_wrap(~ period) +
  labs(fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))

ggsave(plot = plot_spatial_distribution_all_ornaments_by_period,
       here("analysis", "figures",
       "plot-spatial-distribution-all-ornaments-by-period.png"),
       width = 190, 
       height = 50,
       units = "mm")
knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-spatial-distribution-all-ornaments-by-period.png"))
```

```{r, plot-spatial-pattern-each, fig.cap="Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
ornaments_period_join_selected_types <- 
  ornaments_period_join %>% 
  filter(Categories %in% c('Golden bead', 
                           'Glass bead',
                           'Agate bead'))

plot_spatial_distribution_some_ornaments_by_period <- 
ggplot() +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_selected_types, 
          aes(geometry = geometry,
              fill = n), alpha = 0.9) +
  facet_grid(rows = vars(Categories),
             cols = vars(period)) +
  labs(title = "",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 8))

ggsave(plot = plot_spatial_distribution_some_ornaments_by_period,
       here("analysis", "figures",
       "plot-spatial-distribution-some-ornaments-by-period.png"),
       width = 190, 
       height = 130,
       units = "mm")

knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-spatial-distribution-some-ornaments-by-period.png"))
```

@@

```{r point-pattern-data-tidy}
# get centroid and make a list of three periods
ornaments_period_join_cen_all <-
  ornaments_period_join %>% 
  mutate(lon = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[1]])),
         lat = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[2]]))) 
ornaments_period_join_cen <- 
  ornaments_period_join_cen_all %>% 
  group_by(period) %>% 
  group_split() %>% 
  set_names(c("Before European Contact",
              "European Contact",
              "Chinese Contact"))
```
